### 2015-04-01 TODO.Rmd

# To-Do List

* Use permanent e-mails instead of Williams e-mails.