### 2015-04-01 TODO.Rmd
more TODOs

# To-Do List

* Use permanent e-mails instead of Williams e-mails.
* Transfer any undone TODOs from that other system over hear.
* Format vignette as The R Journal submission
* Improve Description in DESCRIPTION file. This material will also be useful in the abstract of the paper.
* Maybe help(package = "qmj") ought to produce something a bit nicer? Not sure about this.
* Set examples to don't run (or something along that line) since they take an absurd amount of time to run and, therefore, make the default running of R CMD check impossible.
* Get rid of lots of useless functions like view_growth and plot_quality. This is overkill. We might use versions of these functions in the article. But we don't need them to be user functions.
* Add Quantitative Value book to bibliography.
* Add some articles from The R Journal about finance backages, including backtest, to the bibliography.