### 2015-04-03 TODO.Rmd
update

# To-Do List

* Format vignette as The R Journal submission (R Markdown)
* Take care of issues raised in R CMD check (format data size)
* 0.2.0 or 0.2-1? What is correct version number format?