### 2015-07-27 TODO.Rmd
minor updates

# To-Do List

* Format vignette as The R Journal submission (R Markdown)
* Take care of issues raised in R CMD check (format data size)